---
title: "Bankruptcy Data"
output:
  html_document:
    df_print: paged
---
This dataset is a subset of Man Xu’s project containing the financial information of the companies. It contains 5436 observations with 13 variables. Variable CUSIP contains unique code to identify the company, variable FR determines the fiscal year and Variable DLRSN is the Bankruptcy/Non Bankruptcy flag, where 1 stands for Bankruptcy while 0 stands for Non Bankrupt companies. Variables R1 to R10 contain financial information which will be used while building a logistic regression model.

```{r}
library(tidyverse)
library(DT)
library(glmnet)
library(rpart)
library(caret)
library(knitr)
```


Bankruptcy dataBankruptcy data (Case) –subset of Man Xu's project (Pages 20-23 for detail). Ten covariates:
R1=Working Capital/Total Asset
R2=Retained Earning/Total Asset
R3=Earning Before Interest & Tax/Total Asset
R4=Market Capital/ Total Liability
R5=SALE/Total Asset
R6=Total Liability/Total Asset
R7=CurrentAsset/Current Liability
R8=Net Income/Total Asset
R9=LOG(SALE)
R10=LOG(Market Cap)

```{r}
data<-read.csv("data/bankruptcy.csv",header=T)
head(data)
```

```{r}
dim(data)
colnames(data)
```
```{r}
data<-data[,-c(1,3)]
head(data)
```


```{r}
str(data)
```
```{r}
colSums(is.na(data))
```

```{r}
#install.packages("ExPanDaR")
library(ExPanDaR)
```

```{r}
#ExPanD(data, export_nb_option = TRUE)
```



```{r}
summary(data)
```

```{r}
mcor<-round(cor(data[,c(2:11)]),2)
mcor
```

```{r}
upper<-mcor
upper[upper.tri(mcor)]<-""
upper<-as.data.frame(upper)
upper
```
```{r}
library(xtable)
print(xtable(upper), type="html")
```


```{r}
library(ggplot2)
ggplot(data,aes(x=data$R1))+geom_histogram(color="black",fill="white")
```

```{r}
library(ggplot2)
ggplot(data,aes(x=data$R2))+geom_histogram(color="black",fill="white")
```

```{r}
library(ggplot2)
ggplot(data,aes(x=data$R4))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R3))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R5))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R6))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R7))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R8))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R9))+geom_histogram(color="black",fill="white")
```

```{r}
ggplot(data,aes(x=data$R10))+geom_histogram(color="black",fill="white")
```


```{r}
boxplot(data[,2:6])
```

```{r}
boxplot(data[,7:11])
```

```{r}
library(psych)
pairs.panels(data[c("R1","R2","R3","R4","R5","R6","R7","R8","R9","R10")],hist.col="green",gap=0)
```

```{r}
library(dplyr)
data<- rename(data,bankruptcy=DLRSN)
```
```{r}
head(data)
```

```{r}
dim(data)
```

```{r}
set.seed(13383610)
index <- sample(nrow(data),nrow(data)*0.70)
data.train = data[index,]
data.test = data[-index,]
```

```{r}
# Logistic regression
data.glm<-glm(bankruptcy~.,family = binomial,data=data.train)
summary(data.glm)
```
```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(jtools)
library(kableExtra)
```

```{r}
summ(data.glm)
```


```{r}
data.glm$deviance

```
```{r}
AIC(data.glm)
```

```{r}
BIC(data.glm)
```
```{r}
hist(predict(data.glm))
```
```{r}
hist(predict(data.glm,type="response"))
```

```{r}
table(predict(data.glm,type="response")>0.5)*1

```

```{r}
table(predict(data.glm,type="response")>0.2)*1
```
```{r}
library(ROCR)
predict2.train<- predict(data.glm,type="response")
pred4 <- prediction(predict2.train, data.train$bankruptcy)
perf4 <- performance(pred4, "tpr", "fpr")
plot(perf4, colorize=TRUE)
```

```{r}
unlist(slot(performance(pred4, "auc"), "y.values"))
```
```{r}
#Binary classificartion
predict.glm.train<-predict(data.glm,type="response")

#convert to binary
pcut1<-mean(data.train$bankruptcy)
pcut1
class.glm.train<- (predict.glm.train>pcut1)*1
```

```{r}
#misclassification rate
MR<-mean(data.train$bankruptcy!=class.glm.train)
#false positive rate
FPR<-sum(data.train$bankruptcy == 0 & class.glm.train == 1)/sum(data.train$bankruptcy==0)
FPR

FNR<-sum(data.train$bankruptcy == 1 & class.glm.train == 0)/sum(data.train$bankruptcy == 1)
MR
```


**Variables Selection using step AIC**
```{r}
fullmodel<- glm(bankruptcy ~ .,
                  data=data.train,
                  family=binomial)

nullmodel = glm(bankruptcy~1,data=data.train,family=binomial)
model_step_both <- step(nullmodel, direction = 'both',
                        scope = list(lower=nullmodel,upper=fullmodel))

summary(model_step_both)
```
```{r}
tab_model(model_step_both)
```

```{r}
summ(model_step_both)
```


```{r}
AIC(model_step_both)
```

```{r}
BIC(model_step_both)
```

```{r}

```


**Variables Selection using step BIC**
```{r}
model_step_both_bic <- step(nullmodel, direction = 'both',
                        scope = list(lower=nullmodel,upper=fullmodel),k=log(nrow(data.train)))

summary(model_step_both_bic)
```

```{r}
tab_model(model_step_both_bic)
```

```{r}
summ(model_step_both_bic)
```


```{r}
BIC(model_step_both_bic)
```
```{r}
AIC(model_step_both_bic)
```
```{r}

```

**ROC of Train Predictions**

```{r}
predict.train<- predict(model_step_both,type="response")
library(ROCR)
pred <- prediction(predict.train, data.train$bankruptcy)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```

```{r}
unlist(slot(performance(pred, "auc"), "y.values"))
```

```{r}
p <- performance(pred,"auc")
as.numeric(p@y.values)
```

```{r}
pred1 <- prediction(predict(model_step_both_bic), data.train$bankruptcy)
perf1 <- performance(pred1, "tpr", "fpr")
plot(perf1, colorize=TRUE)
```

```{r}
unlist(slot(performance(pred1, "auc"), "y.values"))
```

### Testing the out of sample performance: 

**Symmetric Cost**


```{r}
#Binary classificartion
predict.train<-predict(model_step_both,type="response")

#convert to binary
pcut1<-mean(data.train$bankruptcy)
class.train<- (predict.train>pcut1)*1

#evaluate perfromacne
table(data.train$bankruptcy,class.train,dnn=c("True","Predicted"))

```

```{r}
#misclassification rate
MR<-mean(data.train$bankruptcy!=class.train)
MR
```
**Asymmetric** 

```{r}
#dtermine optimal cutoff

# define a cost function with input "obs" being observed response 
# and "pi" being predicted probability, and "pcut" being the threshold.
costfunc = function(obs, pred.p, pcut){
  weight1 = 35   # define the weight for "true=1 but pred=0" (FN)
  weight0 = 1    # define the weight for "true=0 but pred=1" (FP)
  c1 = (obs==1)&(pred.p<pcut)    # count for "true=1 but pred=0"   (FN)
  c0 = (obs==0)&(pred.p>=pcut)   # count for "true=0 but pred=1"   (FP)
  cost = mean(weight1*c1 + weight0*c0)  # misclassification with weight
  return(cost) # you have to return to a value when you write R functions
} # end of the function

```

```{r}
p.seq=seq(0.01,1,0.01)
cost=rep(0,length(p.seq))
for( i in 1:length(p.seq)){
  cost[i]=costfunc(obs=data.train$bankruptcy, pred.p = predict.train ,pcut=p.seq[i])
}


```

```{r}
#find the optimal cut
optimal.pcut.glm = p.seq[which(cost==min(cost))]
optimal.pcut.glm
plot(p.seq, cost)
min(cost)
```


```{r}
predict.test <- predict(model_step_both, newdata=data.test,type = "response")

pred.train <- as.numeric(predict.train > optimal.pcut.glm)
pred.test <- as.numeric(predict.test > optimal.pcut.glm)

```

```{r}
confusion_matrix_train <- table(data.train$bankruptcy, pred.train,dnn=c("True","Predicted"))
confusion_matrix_test <- table(data.test$bankruptcy, pred.test,dnn=c("True","Predicted"))
```

```{r}
confusion_matrix_train
```

```{r}
confusion_matrix_test
```



```{r}
misclassification_rate_train <- round((confusion_matrix_train[2]+confusion_matrix_train[3])/sum(confusion_matrix_train), 2)
misclassification_rate_test <- round((confusion_matrix_test[2]+confusion_matrix_test[3])/sum(confusion_matrix_test), 2)

cat("train misclassfication rate:", misclassification_rate_train, "| test misclassfication rate:", misclassification_rate_test)
```




```{r}
pred2 <- prediction(predict.test, data.test$bankruptcy)
perf2 <- performance(pred2, "tpr", "fpr")
plot(perf2, colorize=TRUE)

```

```{r}
unlist(slot(performance(pred2, "auc"), "y.values"))
```

**Symmetric**

```{r}
#dtermine optimal cutoff

# define a cost function with input "obs" being observed response 
# and "pi" being predicted probability, and "pcut" being the threshold.
costfunc = function(obs, pred.p, pcut){
  weight1 = 1   # define the weight for "true=1 but pred=0" (FN)
  weight0 = 1    # define the weight for "true=0 but pred=1" (FP)
  c1 = (obs==1)&(pred.p<pcut)    # count for "true=1 but pred=0"   (FN)
  c0 = (obs==0)&(pred.p>=pcut)   # count for "true=0 but pred=1"   (FP)
  cost = mean(weight1*c1 + weight0*c0)  # misclassification with weight
  return(cost) # you have to return to a value when you write R functions
} # end of the function

```

```{r}
p.seq=seq(0.01,1,0.01)
cost=rep(0,length(p.seq))
for( i in 1:length(p.seq)){
  cost[i]=costfunc(obs=data.train$bankruptcy, pred.p = predict.train ,pcut=p.seq[i])
}

```

```{r}
#find the optimal cut
optimal.pcut.glm = p.seq[which(cost==min(cost))]
optimal.pcut.glm
plot(p.seq, cost)
min(cost)
```

```{r}
predict.test <- predict(model_step_both, newdata=data.test,type = "response")

pred.train <- as.numeric(predict.train > optimal.pcut.glm)
pred.test <- as.numeric(predict.test > optimal.pcut.glm)
```

```{r}
confusion_matrix_train <- table(data.train$bankruptcy, pred.train,dnn=c("True","Predicted"))
confusion_matrix_test <- table(data.test$bankruptcy, pred.test,dnn=c("True","Predicted"))
```

```{r}
confusion_matrix_train
```

```{r}
confusion_matrix_test
```

```{r}
misclassification_rate_train <- round((confusion_matrix_train[2]+confusion_matrix_train[3])/sum(confusion_matrix_train), 2)
misclassification_rate_test <- round((confusion_matrix_test[2]+confusion_matrix_test[3])/sum(confusion_matrix_test), 2)

cat("train misclassfication rate:", misclassification_rate_train, "| test misclassfication rate:", misclassification_rate_test)
```

```{r}
pred3 <- prediction(predict.test, data.test$bankruptcy)
perf3 <- performance(pred3, "tpr", "fpr")
plot(perf3, colorize=TRUE)
```

```{r}
unlist(slot(performance(pred3, "auc"), "y.values"))
```



